---
title: "Edges in scatterplot displays"
author: "Ursula Laa"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Edges in scatterplot displays}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = TRUE,
  comment = "#>",
  rmarkdown.html_vignette.check_title = FALSE
)
```

```{r setup}
library(tourr)
```

Showing connections between points can be useful in different applications, including geometric shapes, indicating patterns in sequential series or showing connections suggested by a model (for example in clustering). These edges can be added in the 2D scatterplot display (`display_xy`) through the `edges` argument, and should be specified as a two column integer matrix of the indices of ends of lines.

For example, we can generate cube vertices and edges using the `geozoo` package. The information stored as `points` constitutes the input data, and `edges` can directly be passed into the display.

```{r}
# generate 3D cube vertices
cube <- geozoo::cube.iterate(3)
# data is stored points, edges contains the needed two column matrix for connecting points
cube$points
cube$edges
# call grand tour with the scatterplot (xy) display, turning off axes
animate_xy(cube$points, edges = cube$edges, axes = "off")
```

We can use the same functions for higher dimensional cubes as well (but would not print the data).

```{r}
cube5 <- geozoo::cube.iterate(5)
animate_xy(cube5$points, edges = cube5$edges, axes = "off")
```

It can often be useful to connect points with edges, for example sequential points in a time series, or points that are connected by a model. As a simple example we can connect centroids found when clustering the flea dataset:

```{r}
# get centroids of 3 clusters of the flea data
f <- apply(flea[,1:6], 2, function(x) (x-mean(x))/sd(x))
n <- nrow(f)
set.seed(1019)
flea_centroids <- stats::kmeans((f), 3)$centers
flea_aug <- rbind(f, flea_centroids)
col <- c(rep("black", n), rep("orange", 3))
flea_edges <- matrix(c(n+1, n+2, n+1, n+3, n+2, n+3), ncol=2, byrow = TRUE)
animate_xy(flea_aug, edges = flea_edges, 
           col = col, edges.col = "orange", 
           edges.width = 3)
```
