---
title: "global-structure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{global-structure}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(tourr)
library(tidyverse)
library(patchwork)

load(here::here("vignettes", "data", "temp.rda"))
load(here::here("vignettes", "data", "polish_1d.rda"))
load(here::here("vignettes", "data", "polish_2d.rda"))
load(here::here("vignettes", "data", "interrupt_no.rda"))
load(here::here("vignettes", "data", "interrupt_yes.rda"))
```


The latest tour package has implemented a global structure that captures all the information generated during the optimisation routine. This information will be stored and printed when the optimisation ends and can be turned off via `print = FALSE`. Additional messages during the optimisation can be displayed via `verbose = TRUE`.

All the examples below uses simulated data and the histogram of each is presented

```{r echo= FALSE}
set.seed(1234)
x1 <- rnorm(1000, 0, 1)
x2 <- c(rnorm(500, -3, 1), rnorm(500, 3, 1))
x8 <- rnorm(1000, 0, 1)
x9 <- rnorm(1000, 0, 1)
x10 <- rnorm(1000, 0, 1)


data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
                       x9 = x9, x10 = x10) %>% purrr::map_df(scale) %>% as.matrix()

save(data, file = "vignettes/data/data.rda")
```

```{r echo= FALSE}
set.seed(1234)
x1 <- rnorm(1000, 0, 1)
x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
x8 <- rnorm(1000, 0, 1)
x9 <- rnorm(1000, 0, 1)
x10 <- rnorm(1000, 0, 1)


data_mult <- tibble::tibble(x1 = x1, x2 = x2, x3 = x3,
                            x4 = x4, x5 = x5, x6 = x6, x7 = x7,
                            x8 = x8, x9 = x9, x10 = x10) %>% map_df(scale)
save(data_mult, file = "vignettes/data/data_mult.rda")
```

```{r}
origin_dt_bi <- data_mult %>% 
  dplyr::select(-x3) %>% 
  gather(names, values) %>%
  mutate(names = as_factor(names))
  
origin_dt_bi %>%
  ggplot(aes(x = values)) +
  geom_histogram(binwidth = 0.3) +
  geom_density(aes(y = 0.3 * ..count..)) +
  facet_wrap(vars(names), ncol = 3)
```



## Global Structure

Below I presented an example using simulated data. Four variables are simulated via normal distribution and the only informative variable `x2` is simulated via a bi-modal normal distribution centred at -3 and 3. The data has been pre-processed to have variance of 1.  


```{r eval = FALSE}
devtools::load_all()
set.seed(123456)
temp <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_geodesic),
                  rescale = FALSE, verbose = TRUE)

save(temp, file = here::here("vignettes", "data", "temp.rda"))
```

The first five rows of the global structure is printed below. Column `basis` and `index_val` records the basis tried during the optimisation and its corresponding index value using the index function specified in the tour type. `Info` provides metadata on the stage the current basis is in. All the tour have common stage of  `start` and `interpolation` while different searching methods will given different infos

- `serach_geodesic`
  - `direction_search`: a general infinitesimal direction search based on five randomly sampled direction 
  - `best_direction_search`: the direction_search with the highest index value that will be supplied into the line search
  - `line_search`: five randomly sampled points on the geodesic interpolation path given by the best_direction_search
  - `best_line_search`: the basis with the highest index value on the geodesic interpolation path 
  
- `search_better`
  - `random_serach`: a basis generated as a nearby basis of the current. 
  - `new_basis`: a basis that has index value greater than the current basis.
  
`tries` is updated once the generator is called and `loop` is updated on every iteration in the searching process.  `id` shows the row number of each row and makes it easier to plot diagnoses of the optimisation.


```{r}
temp %>% head(5)

```


The final basis can also be extracted as follows:

```{r}
final <- temp %>% filter(info == "interpolation") %>% pull(basis) %>% tail(1)
final[[1]]

temp %>% filter(info == "interpolation") %>% mutate(id = row_number())%>%
  ggplot(aes(x =id, y = index_val)) + geom_line()
```

## Search Polish

A new searching method, search polish, is proposed in the generator to polish the last basis to the absoluate maximum. Search polish samples `n_sample` number of points near the current basis (The searching neighbourhood is controlled by the parameter `polish_alpha`) and picks the one with the largest index value as the current basis for the next iteration. If no basis is found to have a larger index value than the current basis, the neighbourhood parameter `polish_alpha` will shrink by a cooling parameter (`polish_cooling`) and the searching will continue. The polishing procedure will terminate, when any of the four conditions met: 

- The geodesic distance between the current and newly-found basis (`proj_dist(current, target)`) is less than  1e-3; 
- The percentage improvement from the current basis (`(target - cur_index)/cur_index`) is less than 1e-10;
- The searching neighbourhood (`polish_alpha`) is less than 0.001;
- The process reaches the largest number of iteration: `while(try < polish_max_tries)`

Below presents two examples for 1D and 2D projection.

```{r eval = FALSE}
devtools::load_all()
set.seed(123456)
temp_geo <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_geodesic),
                  rescale = FALSE, verbose = TRUE)

last_basis <- temp_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]] 

devtools::load_all()
set.seed(123456)
temp_polish <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_polish),
                  rescale = FALSE, verbose = TRUE, start = last_basis)

#save(temp_polish, file = here::here("vignettes", "data", "temp_polish.rda"))

```

```{r}
# index_val
polish_1d %>% filter(info == "interpolation") %>% pull(index_val) %>% tail(1)
polish_1d %>% filter(info == "polish_best") %>% pull(index_val) %>% tail(1)

# basis
tour_end_1d <- polish_1d %>% filter(info == "interpolation") %>% pull(basis) %>% tail(1)
polish_end_1d <- polish_1d %>% filter(info == "polish_best") %>% pull(basis) %>% tail(1)
theory_1d <- matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1)
proj_dist(tour_end_1d[[1]], theory_1d)
proj_dist(polish_end_1d[[1]], theory_1d)

interp <- polish_1d %>% 
  filter(info %in% c("interpolation", "polish_best")) %>% 
  mutate(id = row_number()) 


p1 <- interp %>% 
  ggplot(aes(x = id, y = index_val)) + 
  geom_line() + 
  geom_point(aes(col = info), alpha = 0.5) + 
  scale_x_continuous(breaks = seq(0, 75, 10))
  #facet_zoom(xy = info == "polish_best",zoom.size = 0.5, ylim = c(0.9375, 0.9385))

p2 <- interp %>% 
  filter(info == "polish_best") %>% 
  ggplot(aes(x = id, y = index_val)) + 
  geom_line() + 
  geom_point(aes(col = info), alpha = 0.5) + 
  scale_color_manual(values = "#00BFC4") + 
  scale_x_continuous(breaks = seq(66, 74, 2))

p1 / p2 + 
  plot_layout(ncol = 2, widths = c(2, 1), guides = "collect")
  
```



```{r eval = FALSE}
devtools::load_all()
set.seed(123456)
polish_2d <- animate_xy(data_mult[,c(1,2, 7:10)], tour_path = guided_tour(holes(), d = 2,
                                           search_f =  search_geodesic),
                  rescale = FALSE, verbose = TRUE, polish = TRUE)

save(polish_2d, file = here::here("vignettes", "data", "polish_2d.rda"))
```

```{r}
# index_Val
polish_2d %>% filter(info == "interpolation") %>% pull(index_val) %>% tail(1)
polish_2d %>% filter(info == "polish_best") %>% pull(index_val) %>% tail(1)

# basis
tour_end_2d <- polish_2d %>% filter(info == "interpolation") %>% pull(basis) %>% tail(1)
polish_end_2d <- polish_2d %>% filter(info == "polish_best") %>% pull(basis) %>% tail(1)
theory_2d <- matrix(c(0, 1, 0, 0, 0, 0, 
                      0, 0, 1, 0, 0, 0 ), nrow = 6, ncol = 2)
proj_dist(tour_end_2d[[1]], theory_2d)
proj_dist(polish_end_2d[[1]], theory_2d)

interp <- polish_2d %>% 
  filter(info %in% c("interpolation", "polish_best")) %>% 
  mutate(id = row_number()) 


p1 <- interp %>% 
  ggplot(aes(x = id, y = index_val)) + 
  geom_line() +
  geom_point(aes(col = info), alpha = 0.5)
  #facet_zoom(xy = info == "polish_best",zoom.size = 0.5, ylim = c(0.9375, 0.9385))

p2 <- interp %>% 
  filter(info == "polish_best") %>% 
  ggplot(aes(x = id, y = index_val)) + 
  geom_line() + 
  geom_point(aes(col = info), alpha = 0.5) + 
  scale_color_manual(values = "#00BFC4")

p1 / p2 + 
  plot_layout(ncol = 2, widths = c(2, 1), guides = "collect")
  
```


## Interruption

In search_better, it is likely that a basis with even higher index value can be found on the interpolation path from the current to the target basis. This is shown as a "hill" shape on the tracing plot. An interruption is implemented to allows the next iteration to start from the basis with the highest index value rather than from the ending basis. With this interruption, the search function is likely to find basis with higher index_val in the end. 

```{r eval =FALSE}
devtools::load_all()
set.seed(123456)
interrupt_no <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_better),
                  rescale = FALSE, verbose = TRUE)

save(interrupt_no, file = here::here("vignettes", "data", "interrupt_no.rda"))

devtools::load_all()
set.seed(123456)
interrupt_yes <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_better),
                  rescale = FALSE, verbose = TRUE)
save(interrupt_yes, file = here::here("vignettes", "data", "interrupt_yes.rda"))

```


```{r}
interrupt_no %>% filter(info == "interpolation") %>% pull(index_val) %>% tail(1)
interrupt_yes %>% filter(info == "interpolation") %>% pull(index_val) %>% tail(1)

p1 <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>%
  ggplot(aes(x =id, y = index_val)) + geom_line() + 
  ggtitle("no interruption")

p2 <- interrupt_yes %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>%
  ggplot(aes(x =id, y = index_val)) + geom_line() + 
  ggtitle("with interruption")

p1 / p2 + 
  plot_layout(ncol = 2)
```

