---
title: "Using the skinny index with jellyfish search"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the skinny index with jellyfish search}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  eval = TRUE,
  rmarkdown.html_vignette.check_title = FALSE
)
```

The `skinny` scagnostic basically tells you how thin a 2d point cloud looks.
It comes from the `cassowaryr` package and is based on the ratio of the
alpha hull perimeter to its area — so higher values mean the points are stretched
out in a narrow, elongated way.

Here we'll try using `skinny()` as a projection pursuit index with `guided_tour`,
and use `search_jellyfish` as the optimiser since it tends to explore more broadly.

```{r setup, message = FALSE}
library(tourr)
```

## Setting up the data

We'll work with the `flea` data that comes with tourr. Just standardising the
numeric columns first so they're on a comparable scale.

```{r}
f <- apply(flea[, 1:6], 2, function(x) (x - mean(x)) / sd(x))
head(f)
```

## Guided tour with skinny

We can pass `skinny()` into `guided_tour()` the same way you'd use `holes()`
or any other index. Adding `search_f = search_jellyfish` for the optimiser:

```{r}
animate_xy(f,
  tour_path = guided_tour(skinny(),
    search_f = search_jellyfish),
  col = flea$species)
```

If you run this in the console you'll see it animate — the tour tries to
find projections where the data looks as thin/elongated as possible.

## Saving the tour path

It can be handy to save the tour and replay it afterwards, especially
if you want to look at it in a different display or just keep a record.

```{r}
set.seed(2025)
path <- save_history(f,
  guided_tour(skinny(), search_f = search_jellyfish),
  max = 10)
animate_xy(f, planned_tour(path), col = flea$species)
```

## Quick comparison with holes

Just to see how `skinny()` differs from something like `holes()` — the holes
index looks for projections with low density in the middle (good for finding
clusters), while skinny is looking for elongated structure.

```{r}
set.seed(2025)
path_h <- save_history(f, guided_tour(holes()), max = 10)
animate_xy(f, planned_tour(path_h), col = flea$species)
```

You can see that `holes()` tends to tease apart the species groups, while
`skinny()` finds a different kind of structure — more stretched out views
of the data.
