---
title: "Guided tour with the skinny scagnostic index"
author: "Vaibhav Manihar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guided tour with the skinny scagnostic index}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  eval = TRUE,
  rmarkdown.html_vignette.check_title = FALSE
)
```

```{r setup}
library(tourr)
```

The `skinny()` index is a scagnostic measure computed using the `cassowaryr` package. It measures how narrow or elongated a point cloud appears in a 2D projection, and is useful for finding projections that reveal linear or elongated structure in the data.

## Using skinny() in a guided tour

We use the `flea` dataset, standardising the numeric variables before running the tour.

```{r}
f <- apply(flea[, 1:6], 2, function(x) (x - mean(x)) / sd(x))
```

We can inspect the index value on a specific projection to understand what the scalar output looks like.

```{r}
set.seed(2026)
proj <- basis_random(6, 2)
skinny()(f %*% proj)
```

The guided tour searches for projections that maximise this value, moving towards more elongated views of the point cloud.

```{r}
set.seed(2026)
animate_xy(f,
  tour_path = guided_tour(skinny()),
  display = display_xy(col = flea$species))
```

## Using the jellyfish optimiser

The jellyfish optimiser (`search_jellyfish`) is a population-based search that can escape local optima more effectively than the default search. The tour path can be saved and replayed as a planned tour.

```{r}
set.seed(1234)
path <- save_history(f,
  guided_tour(skinny(), search_f = search_jellyfish),
  max = 10)
animate_xy(f, planned_tour(path), col = flea$species)
```

## Comparison with holes()

The `holes()` index finds projections with low density near the centre, which tends to reveal cluster separation. The `skinny()` index instead finds elongated structure, producing qualitatively different views of the same data.

```{r}
set.seed(1234)
path_holes <- save_history(f, guided_tour(holes()), max = 10)
animate_xy(f, planned_tour(path_holes), col = flea$species)
```
