---
title: "tourr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tourr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tourr)
verbose <- F
```

A tour animates a display of interpolated low-dimensional projections from high-dimensional data, to explore the shape of the distribution. This vignette shows how to use the `tourr` package to generate animations using different tour and display types.

## Quick start

To get started we can simply generate a tour animation for any numeric data matrix. Here we look at the flea data available in the package, drop the last column (the grouping variable), and call

```{r}
f <- flea[, 1:6]
animate(f)
```

Here we only see the first frame that is generated on the tour path, but if you call `animate` in your console you will launch a tour animation that you can view in your graphics window.

We could have also launched the same tour by specifying the default selection explicitly

```{r}
animate(f,
        tour_path = grand_tour(),
        display = display_xy()
        )
```

The most common application is to use a grand tour, which means that the projections (or bases) are randomly selected and give a global overview of the distribution. The typical display is a scatterplot display, i.e. working with 2-dimensional projections.

## Tour types

As we have seen, the default tour type is a grand tour. By selecting the second argument, `tour_path`, in the animate function we can change this, choosing from the following tour types that have been implemented in the package. The most relevant options are:

- grand tour: randomly selected bases
- guided tour: basis selection is optimising an index function
- planned tour: interpolates between a set of input bases
- little tour: a planned tour between all axis parallel projections
- local tour: alternating between the starting projection and randomly selected nearby projections

The guided tour is aiming to move towards more interesting views of the distribution as the animation progresses. To measure the interestingness of each projection we need to define an index function which will get maximised along the tour path. You can define your own index function, or use one of those available in the `tourr` package. For our example flea data we can use the holes index which is looking for projections with low densities near the center, and can often find views that reveal clustering.

```{r}
animate(f,
        tour_path = guided_tour(holes()),
        display = display_xy())
```


## Display types

The default display is showing a scatterplot of the data projected to 2D. Depending on the number of dimensions we project onto (this is called d), we can choose different display types. For example, we could look at 1D projections in a density display, a 3D projection using depth cues, or higher dimensional projections in a parallel coordinate or scatterplot matrix display.

When changing the display type via the `display` argument, we need to make sure that the basis generation is matching the corresponding number of dimensions d. We can pass in this infromation when generating the tour path. To work with 1D projections and a density display, we can launch the animation as

```{r}
animate(f,
        tour_path = grand_tour(d = 1),
        display = display_dist()
        )
```

but there is also a shortcut available

```{r}
animate_dist(f)
```

## Saving a tour

After exploring the different options, we may have identified a particularly interesting tour that we may want to save or share.

The first option is to save the animation (or its individual frames). This is possible through the `render` functions, that save the frame views to png or pdf format. Another option, perhaps more convenient, is to directly save the full animation to a gif file. To use the function, you will need to install the `gifski` package.

An advanced alternative is to save the tour path, which can later be replayed as a planned tour, and allows us to look at the same tour in different displays. For example we can save a default grand tour path for the flea data and then replay it in a scatterplot display

```{r}
t1 <- save_history(f, max = 3)
animate(f, planned_tour(t1))
```

